Standard Candle Cosmology With Incomplete Spectroscopic Classification

### 2015-10-22 doc/article.tex
new version

%SNe~Ia are not perfect standard candles but rather are calibrated candles whose
%peak brightnesses are a
%function of their internal properties discerned from their colors and light-curve evolution.
%Measurement uncertainty, particularly from photometric flux
%calibration and galaxy photometric redshifts, can play an important role in the statistical
%reach of experiments. 
The notation for the likelihood presents the data and fundamental model parameters
\begin{multline}
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, p(n, {{T}}_S,{{z}}_S, \theta_{S}, G | T, \theta_{Ia},
\theta_{!Ia}, L, \mathcal{S}, \Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, g) \\
p(L|T, \theta_{Ia},
\theta_{!Ia},  \mathcal{S}, \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g)
p(T, \theta_{Ia},
\theta_{!Ia}|  \mathcal{S}, \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g),
\label{likelihood2:eqn}
\end{multline}
whose relations are sketched in the probabilistic graphical model in Figure~\ref{pgm:fig}.
$\mathcal{S}$ indicates that this likelihood corresponds to the analysis sample. 
For shorthand we asign $A=\{n, {{T}}_S,{{z}}_S, \theta_{S}, G \}$, $B=\{T, \theta_{Ia},
\theta_{!Ia}, L\}$, and $C=\{\Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, \}$
so that the likelihood is more compactly expressed as $\int dB\, p(AB| \mathcal{S}, C)p(B| \mathcal{S}, C)$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{/Users/akim/project/abc/results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from left to right, starting with the description of the universe, observatory,
   and data.    A transient has host galaxy $g_i$, which determines its redshift $z_i$
   and galaxy parameters $\theta_{gi}$.
   A model with parameters $\theta_\mu$ fix the distance modulus $\mu_i$.
   The transient type $T$ depends on the host-galaxy parameters  $\theta_{gi}$
   and rates $r$.   The transient
   parameters $\theta_T^X$, $\theta_{Ti}^X$ (and perhaps the host galaxy) determine the luminosity $L$.       The 
   incoming photon flux $n$, $n_g$  are then fixed
   with redshift and distance modulus.
   The instrumental transmission function $\phi$ is calibrated with data ${Z}$ andflux
   gives the expected
   counts $\overline{f}$, $\overline{f}_g$. 
   The transient has realized light curves ($n$) and passes selection criteria
   through $\tau_i$ to be included in the sample.  Some transients have spectral data
   ${X}_S$.  The properties of the galaxy identified as the host galaxy is $G$. 
   \label{pgm:fig}}
\end{figure}

The measurements are
\begin{itemize}
\item $n$: the measured counts for each photometric observation for all transients in the sample;
\item ${{T}}_S$, ${{z}}_S$, $\theta_S$: the spectroscopic type, redshift, and transient parameters available for a subset of transients;
\item $G$: measured properties of the galaxy identified
as the host.
\end{itemize}
The top-level model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the mass density and dark-energy equation of state parameters of a flat,
constant-$w$ dark-energy cosmology;
\item $r$:  the parameters
that describe the redshift-dependent relative fraction of SNe~Ia and non-Ia's;
\item $\Theta_{Ia}$, $\Theta_{\mathit{!Ia}}$:  All the fundamental parameters used to describe
the SN~Ia and non-Ia populations respectively, such as the mean
absolute magnitude, intrinsic magnitude dispersion, the distribution of internal parameters
(i,e,\ the distributions of SALT2's $x_1$ and $c$), the relationship between internal
parameters and magnitude (i.e.\ SALT2's $\alpha$, $\beta$);
\item $Z$: the parameters that describe the transmission functions, i.e.\ photometric calibration;
\item $g$: the parameters that describe the candidate host and neighbor galaxies
including redshift $g_z$.
\end{itemize}
The second-level parameters are
\begin{itemize}
%\item $z$: the cosmological redshifts of the host and neighboring galaxies;
\item $T$: the class of the objects, SN~Ia or non-Ia;
\item $\theta_{Ia}$, $\theta_{!Ia}$: the transient parameters that describe individual
transients (i.e..\ SALT2's $x_1$, $c$);
\item $L(t,\lambda)$: the luminosity of the transient.
\end{itemize}
There are  fixed parameters
\begin{itemize}
\item $\mu$: the distance modulus given the cosmological parameters and the redshift;
\item{$f$}: the flux of the transient given the luminosity, distance modulus, and
redshift.
%\item{$\overline{n}$}: the expected counts of the transient given
%the flux and transmission.
\end{itemize}

The relationships between the parameters are now described:
\begin{itemize}
\item Intrinsic transient parameters are dependent on their distributions and the host galaxy.
The SNe~Ia parameters are drawn as
\begin{equation}
\theta_{Ia} | \Theta_{Ia}, g \sim \mathcal{N}(0, \Theta_{ia, \theta}),
\end{equation}
where $\theta_{Ia} =\{x_1, c\}$ are the SALT2 light-curve shape and color parameters
and $\Theta_{ia, \theta}$ describes the covariance of their distributions.
A more complex model can be based on non-Normal distributions
and/or including SN properties directly tied to characteristics (e.g.\ mass, star formation
rate) of the host galaxy.

{\bf The non-Ia parameters, their underlying distributions, and SED's need to chosen.}
The prime candidate
is their description used in PSNID.

%\begin{equation}
%\theta_{non-Ia} | \sigma_\beta, g  \sim \mathcal{N}(0, \sigma_{\beta, non-Ia}),
%\end{equation}
%where the parameter is the realization of an intrinsic magnitude dispersion.
%A more complex model can include multiple non-Ia populations each with its own
%intrinsic parameters.
\item Spectroscopically observed transient parameters depend on the intrinsic transient
parameters and type.  $\theta_{Si}$ may be ignored in a zero'th order analysis that
does not relate spectral features with SN~Ia absolute magnitudes.   If spectral
features are considered, the simple model
\begin{equation}
\theta_{S} (\theta_{Ia}, \theta_{!Ia}, T)=\theta_{T}
\end{equation}
assumes an unambiguous mapping between supernova parameters and spectral signal.
{\bf We need to decide what SN spectral parameters, if any, we will consider in our analysis.}
\item The type is drawn from the host-dependent relative rate between SNe~Ia and non-Ia's 
\begin{equation}
T | r, g \sim \rm{Bernoulli}(\pi(r,g,T)),
\end{equation}
where $\pi$ is a model for the relative rate of SNe~Ia as a
function of galaxy parameters and redshift.  {\bf A reasonable choice for $\pi$ needs to be
selected.}
\item Supernova type is assumed to be unambiguously encoded in the
spectrum:
\begin{equation}
T_S (T)  = T.
\end{equation}
\item The observed redshift is taken to be the host galaxy redshift
\begin{equation}
z_S (g_z) =g_z
\end{equation}
where $g_z$ is the element in $g$ that corresponds to redshift.
Added sophistication includes redshift
measurement uncertainty and random peculiar velocities. The study of coherent
flow requires added model complexity.
\item The underlying SED of transients are modeled.  For SNe~Ia the SALT2 model
is an option, although new models that separate dust and intrinsic color or
identify new SN~Ia parameters are expected in the near future.  Absolute magnitudes
that depend on host-galaxy properties need to be considered. 
\begin{equation}
L(t,\lambda) |  \Theta_{Ia}, \theta_{Ia}, g \sim L_{SALT2}(t-t_0; \alpha, \beta, x_1, c) \gamma(g, \Theta_{Ia,g}) \mathcal{N}(0,\sigma_{Ia}),
\end{equation}
where the first  term is the SALT2 light-curve and absolute magnitude
model with a realized error snake,
and $\gamma(g, \Theta_{Ia})$ a step function to model the ``mass step''.  Recall that the  fundamental SALT2 parameters 
$\alpha$, $\beta$, and intrinsic dispersion $\sigma_{Ia}$, and the parameters that
relate SN magnitude and host,
are included in $\Theta_{Ia}$; the parameters in the SALT2 fit $t_0$, $x_1$, $c$
are included in $\theta_{Ia}$.

{\bf The non-Ia model needs to be identified.  Perhaps PSNID.}
\item The theoretical distance modulus
\begin{equation}
\mu(g_z; \Omega_M, w) = \mu_{wCDM}(g_z; \Omega_M, w)
\end{equation}
is the predicted distance modulus of a flat dark energy cosmology.
\item The flux incident on Earth is the redshifted and time-dilated SED scaled
by the distance modulus.
\begin{equation}
f(t,\lambda; L) = 10^{-\mu/2.5} L\left(t/(1+g_z),\lambda/(1+g_z)\right)
\end{equation}
\item The expected light curve in counts depends on the flux and throughput of the
optical chain.  A first model encapsulates flux calibration through a zeropoint
\begin{equation}
n(f, Z) = 10^{Z/2.5} \int  f(\lambda) T_0(\lambda)d\lambda,
\end{equation}
where $T_0$ is a fiducial transmission function for the appropriate filter.  A more sophisticated
analysis can have a transmission function specific to each observation.
\item The host-galaxy information is directly encoded in the measurements
\begin{equation}
G(g)=g.
\end{equation}
\end{itemize}

\subsection{Measurements}
\label{measurements:sec}
Spectroscopic measurements are assumed
to provide prefect determinations of transient type and parameters, and
galaxy redshifts and parameters. 
\begin{align}
p(T_S|T) &= \delta(T_S-T),\\
p(\theta_S|\theta_T) &= \delta(\theta_S-\theta_T),\\
p(z_S|g_z) &= \delta(z_S-g_z).
\label{specz:eqn}
\end{align}
People who are interested in SN spectral features
or host-galaxy correlations (Chris, Rachel, Mat) can develop these models.

For the photometric sample, there is uncertainty in the identity of the host galaxy.
From a set of potential host galaxies $\hat{G}$ each with probability $\pi$ of being the correct host,
\begin{equation}
G \sim \prod_i \delta(G-\hat{G}_i) \pi_i.
\end{equation}
Ravi can provide for each transient a list of candidate hosts and their probabilities.
The working group is striving to get spectroscopic
redshifts of all potential hosts, which is included in $G$.
For the spectroscopic sample the host is unambigous
\begin{equation}
G = g.
\end{equation}

Practically the photometric extraction is handled by independent software, which calculates
the $n$ with uncertainties $C_P$
by modeling and separating out polluting background light from the transient
signal
\begin{equation}
n \sim \mathcal{N}\left(10^{Z/2.5} \int  f(\lambda;A, B) T_0(\lambda)d\lambda,C_P\right).
\end{equation}
Dillon is providing this.
Note that the model has explicit calibration parameters that are to be fit.  An alternative
approach is to include calibration uncertainties in $C_P$.

\subsection{Sample Selection}
The transients that enter the analysis satisfy some selection criteria.  Trivially, they
have to be discovered in the first place and often more stringent photometric signal-to-noise
requirements are applied.  Additionally, the probability for a transient to have successful
spectroscopy may depend on parameters.  These criteria have not yet been determined/selected
(Argonne for photometry, OzDES/Ryan for spectroscopy are candidates to do this);
presented here is a sketch of the likelihood calculation for a reasonable
photometric selection criterion.
For each transient $\nu$, the criterion is to have $\ge N$ photometric points
above a signal-to-noise threshold there  given $M_\nu$ available measurements:
the signal-to-noise threshold corresponding to a flux threshold
 $n_{0,\nu i}$.

Call the likelihood of the underlying distribution without sample selection $P(A|C)$.
Let the array $b_{\nu}$ be an array that contains $0$ or $1$ based on
whether the corresponding photometric point is below or passes threshold. 
Let the larger array $b$ be a combination of the $b_\nu$ from all transients.
There are $2^{\sum_\nu M_\nu}$ possible permutations of $b$, and the underlying
likelihood can be broken up into the sum of all possible combinations
\begin{align}
p(A|C) &= \sum_k p(A, b_k| C)\\
& = \sum_k \epsilon(n, b_k)p(A| C),
\end{align}
where
\begin{equation}
\epsilon(n; b_{k})= \prod_{\nu i}\mathcal{H}\left((-1)^{b_{k\nu i}+1}(n_{k\nu i}-n_{0,\nu i})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{k\nu i}=1/0$ the step in the efficiency
is up/down.  The indices correspond to configuration $k$,
transient $\nu$, and photometric point $i$.

For the analysis sample, are a smaller number of combinations of $b$ that satisfy
the selection criteria;
instead of $2^{M_\nu}$ there are $\sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}$ 
distinct $b_{\nu}$'s that allow acceptance of transient $\nu$.  The selection efficiency
for the complement set of possible $b$'s is zero.
The sample likelihood is then as
\begin{equation}
p(A|\mathcal{S}, C) = \frac{\sum_k \epsilon(n, b_k)p(A| C)}{\int dA\, \sum_k \epsilon(n, b_k)p(A| C)},
\end{equation}
where the sum is
over only the configurations that meet the selection criteria.

The measurement uncertainties in the light curves $n$ are independent of the other
measurements, and the measurement uncertainties are drawn from a Normal distribution:
the likelihood is assembled from truncated Gaussians with the integrals over $n$ in the numerator
proportional to error functions.

@@

\subsubsection{Likelihood for the {\it obs} Set}
The likelihood for the {\it obs} set is
\begin{multline}
p(A|\mathcal{S},C) \propto
\sum_k \epsilon(n, b_k) 
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, p(n, {{T}}_S,{{z}}_S, \theta_{S}, G | T,  L, \Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, g) \\
p(L| T, \theta_{Ia},
\theta_{!Ia},  \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g)
p(T, \theta_{Ia}
\theta_{!Ia}|  \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g).
\end{multline}

For our model and measurements
\begin{multline}
p(A|\mathcal{S},C) \propto 
\sum_k \epsilon(n, b_k) \int dL\,
p(n |  T= T_S, \theta_{T}=\theta_S, g=G, L, C)
p(L| T, \theta_{T_S}, r, \Theta_{T_S}, g)\\
\pi(r, z_S, T_S)p( \theta_{T_S}|  \Theta_{T_S},g=G) .
\end{multline}

\subsubsection{Likelihood for the {\it mis} Set}
The likelihood for the {\it mis} set follows the same pattern as that of the
{\it obs} set except for not having to account for spectroscopic data
\begin{multline}
p(n, \theta_{G} |  \mathcal{S}, C) 
 \propto \sum_k \epsilon(n; b_{k}) 
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, 
\sum_l
\pi_l
p(n |  g=\hat{G}_l, B, C)
p(L| T, \theta_{Ia}, \theta_{!Ia}, r, \Theta_{Ia}, \Theta_{!Ia}, g)\\
\pi(r, \hat{G}_l, T)p( \theta_{Ia},
\theta_{!Ia}|  \Theta_{Ia},\Theta_{\mathit{!Ia}}, g=\hat{G}_l).
\end{multline}

%\begin{multline}
%p(n_{SN}, G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
%= \sum_i \int_L dL\, p(n_{SN}, G | b_i, L, \ldots)
%P(b_{i} |L, \ldots) P(L|\ldots) ,
%\end{multline} 
%\begin{equation}
%p(n_{SN}, G | b_i, L, \ldots)
%  =  \frac{ \prod_j
% \epsilon(n_{SN,j}; b_{ij})
%  p(n_{SN}, G | L, \ldots)}
%  {\prod_j P(b_{ij}{ |L,\dots)}}.
%\label{misconf:eqn}
%\end{equation}

>>>>

The Probabilistic Graphical Model is shown in Figure~\ref{pgm:fig}. The underlying behavior of the universe is described by the following elements:
\begin{description}
\item [Cosmology] Luminosity distance for the cosmological model being considered,
e.g.\ a $wCDM$ universe parameterized by $\Omega_M$, $\Omega_w$, $w_0$, $w_a$.
\item[SNe~Ia, Non-Ia Populations] Time-dependent luminosity for all transient populations
that may enter
the analysis.  The SALT2 model is a candidate for SNe~Ia, with parameters $\alpha$, $\beta$,
and pdf's for $x_1$ and $c$.  The PSNID treatment of non-Ia's is a candidate for the other
transients.
\item [Relative Rates] Relative rates of all the transients populations under consideration.  The
model could be extended to measure absolute rates, but that is unnecessary if
the cosmological parameters are the goal.
\end{description}
Each transient is described by the following elements:
\begin{description}
\item[Host Galaxy] The host galaxy, or more generally for the case of hostless supernovae,
the environment from which the transient arises.  The host is described by its redshift
and parameters that may be correlated with Hubble diagram residuals, such as mass,
specific star formation rate, metallicity.
\item[Type Subtype] Specification of the the transient: which population to which it belongs
and its internal parameters.  For example, an object could be SN~Ia with specific
values for its SALT2 parameters $c$, $x_1$, $t_0$ and its realization
from the error snake.  This dependent on the relative rates and
its host galaxy including redshift.
\item[Distance] Distance modulus or equivalently the luminosity distance of the host galaxy,
which in our model corresponds to the cosmological redshift of the transient.  The value
is fixed based on the cosmological model and redshift.
\end{description}
The individual photometric measurements of a transient are described by the following
elements:
\begin{description}
\item[Luminosity]  Luminosity as a function of wavelength
corresponding to the date of observation.   This depends on the population, type,
and subtype.  Given evidence for correlations between host properties and Hubble
residuals, a direction connection between host galaxy and luminosity is made.
\item[Flux] Flux of the transient incident to earth at the date of observation.  It is 
fixed based on the luminosity, redshift, and distance.
\end{description}
The observatory is described by the following element:
\begin{description}
\item[Throughput] The transmission function for the band at that date appropriate
for the transient.  A simple parameterization could be calibration zeropoints; more
sophisticated descriptions can vary the shape, not only the normalization, of the
transmission.
\end{description}
The data subscripted by ``o'' are
\begin{description}
\item[$\text{Photometry}_o$] Photometric flux measurement.  The expected measurement
is the incident flux integrated over the throughput.  Typically the measurement
uncertainties are from a Poisson distribution.  In practice, photometric data extraction
is performed in an independent analysis that isolates the transient flux
from contaminating background light. 
\item[$\text{Type Subtype}_o$] Measurement of the intrinsic transient type.  This is done
through spectroscopy.
\item[$\text{Redshift}_o$] Measurement of the host redshift.  The redshift can be
spectroscopic or photometric.  Uncertain host galaxy identification contribute
to measurement errors.
\item[$\text{Host Properties}_o$] Measurement of the host properties such
as mass, sSFR, metallicity. Uncertain host galaxy identification contribute
to measurement errors.
\end{description}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{/Users/akim/project/abc/results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from top to bottom, starting with the description of the universe, observatory,
   and data.    A transient occurs in host galaxy $g_i$.  The type $T$ depends on the
   host and relative rates $r$.  The transient parameters $\theta$ depend on the
   transient model $\Theta$ and the host.  The luminosity $L$ depends on the preceding
   parameters.  
   A model with parameters $\Omega_M$ and $w$ fix the distance modulus $\mu_i$.
   The 
   incoming photon flux $f$  are then fixed
   with host redshift and distance modulus.
   The measured light curve depends on the instrumental throughput $Z$.  There are
   spectroscopic
   measurements $\theta_S$, $T_S$, $z_S$ of the transient parameters, type, and
   redshift.  $G$ is the determination of the host galaxy from photometric data.
   \label{pgm:fig}}
\end{figure}

This hierarchical model has several features worth noting.  Transients with and without
spectroscopic classification are treated together: those with classification help constrain
the properties of the underlying population to inform the treatment of those without.
Photometric classification is treated internally, through the
latent Type.  The system throughput is treated as a probabilistic variable
allowing for self-calibration; this could be omitted from the model with calibration applied
to the Photometry measurements.  Host-galaxy misclassification enters as measurement
uncertainties.  Malmquist bias and other selection effects do not appear explicitly, but are accounted
for in the likelihoods.  The model could be extended to include summary statistics,
for example the results of an external supernova classifier; the use
of summary statistics would affect the manner in which the model is evaluated
due to our inability to write their likelihoods.

\subsection{Sample Selection}
The transients that enter the analysis satisfy some selection criteria.  Trivially, they
have to be discovered in the first place and often more stringent photometric signal-to-noise
requirements are applied.  Additionally, the probability for a transient to have successful
spectroscopy may depend on parameters.
These criteria have not yet been determined/selected
(Argonne for photometry, OzDES/Ryan for spectroscopy are candidates to do this);
presented here is a sketch of the likelihood calculation for a reasonable
photometric selection criterion.

For each transient $\nu$, the criterion is to have $\ge N$ photometric points
above a signal-to-noise threshold given $M_\nu$ available measurements:
every photometric point is associated with a flux threshold that it either passes
or fails.  Given a total $M=\sum_\nu M_\nu$ data points, there are
$2^M$ possible combinations
of passing or failing theshold, of which only
$\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}$  combinations would lead to
the inclusion of the transients in the sample.

The probability distribution function for all possible data realizations can be written
as the sum of the pdf's of all possible permutations
\begin{equation}
p(data|model) = \sum_{k=1}^{2^M} p(data, b_k| model),
\label{likelihood_u:eqn}
\end{equation}
where $b$ contains all possible combinations.
The probability distribution function for all possible data realizations that pass
our selection criteria is a conditional pdf. It is proportional to the underlying
pdf after removal of conditions that are not achieved, i.e.\ removal from the sum
in Eqn.~\ref{likelihood_u:eqn} of all combinations that do not meet the sample
selection criteria. This leaves a reduced set $b$ of
possible permutations
\begin{equation}
\mathcal{L}(data|\mathcal{S}, model) \propto \sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} p(data, b_k| model),
\end{equation}
where $\mathcal{S}$ is the sample-selection condition.

The probability for the data to be in a specific configuration $b_k$ of
being below or above flux threshold is expressed
as
\begin{equation}
p(data, b_k| model)  = \epsilon(Photometry_o, b_k)p(data| model)\\
\end{equation}
where
\begin{equation}
\epsilon(Photometry_o; b_{k})= \prod_{\nu i}\mathcal{H}\left((-1)^{b_{k\nu i}+1}(Photometry_{o,k\nu i}-Threshold_{\nu i})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{k\nu i}=1/0$ the step in the efficiency
is up/down.  The indices correspond to configuration $k$,
transient $\nu$, and photometric point $i$.

Assembling this together, the likelihood is
\begin{equation}
\mathcal{L}(data|\mathcal{S}, model) = \frac{\sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} \epsilon(Photometry_o, b_k)p(data| model)}{\int ddata\,
\sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} \epsilon(Photometry_o, b_k)p(data| model)
}
\end{equation}
where the normalizing factor is included explicitly as a reminder that it has 
parameter dependence and must not be neglected in calculations.

Consider the traditional case of Malmquist bias of a single object,
where a single measurement has
to be above flux threshold.  For the underlying distribution with no flux threshold there
are $2^1=2$ possible configurations, the flux is either above or below threshold.  For
the sample distribution, there is $\binom{1}{1}=1$ possible configuration: the flux
is above threshold.   The likelihood for the measured flux is the truncated version
of the underlying distribution.

For the case of a single object with two measurements and a single threshold requirement,
there are $2^2=4$ possible configurations of which $\sum_{k=1}^{2} \binom{2}{k}=3$
would allow sample selection, the exception being when both measurements fall below threshold.
The likelihood for each flux is composed from
the sum of three truncated distributions of which one accepts faint fluxes,
resulting in resultant likelihoods that do not fall to zero below detection threshold.

\subsection{Underlying Likelihood}
The likelihood of the analysis sample is calculated from the underlying pdf of data.  Based on Figure~\ref{pgm:fig},
\begin{align}
p(data|model) &=\int dType\, dLuminosity\, p(Photometry_o |  Luminosity, Cosmology,Host,Throughput)  \nonumber\\
& p(Luminosity| Type
, Host, Population) p(Type | Rates, Host)  \nonumber\\
& \times \int dType\, p(Type_0|Type) p(Type|Rates, Host) \nonumber \\
& \times p(Redshift_o, Host_o| Host)
\label{like:eqn}
\end{align}
where $data=\{Photometry_o, Type_o, Redshift_o, Host_o\}$,
$model=\{Cosmology, Populations, Rates\}$, and we anticipate the independence
of photometric and spectroscopic typing measurements.

Now for specificity candidate models for the pdf's that enter Eqn.~\ref{like:eqn}
are presented.
\begin{itemize}
\item The type is drawn from the host-dependent relative rate between SNe~Ia and non-Ia's 
\begin{equation}
Type | Rates, Host \sim {\rm Bernoulli}(\pi(z;r)),
\end{equation}
where $\pi$ is a model for the relative rate of SNe~Ia as a
function of galaxy redshift $z$.  {\bf A reasonable choice for $\pi$ needs to be
selected.}

Type~Ia supernovae have values of SALT2 parameters
$x_1$ and $c$ drawn from Normal distributions
with parameters for the variance
\begin{equation}
x_1, c \sim \mathcal{N}(0, (\sigma_{x_1}, \sigma_c)).
\end{equation}

Models for non-Ia supernovae need to be selected.  The model used in PSNID is a good
place to start.
\item The underlying SED of transients.  For SNe~Ia the SALT2 model
is an option, although new models that separate dust and intrinsic color or
identify new SN~Ia parameters are expected in the near future.  Absolute magnitudes
that depend on host-galaxy properties need to be considered. 
\begin{equation}
Luminosity| Type
, Host, Population \sim L_{SALT2}\left(\frac{t-t_0}{1+z}; \alpha, \beta, x_1, c\right) \exp{\left(\delta_G\mathcal{H}(g)\right)},
\end{equation}
where the first  term is the SALT2 light-curve and absolute magnitude
model with error snake.  Possible host-galaxy effects like the ``mass step'' are
accounted in $\mathcal{H}(g)$ where $g$ is the corresponding host parameter
and $\delta_G$ is the size of the offset.  The host redshift is $z$.

{\bf The non-Ia model needs to be identified.  Perhaps PSNID.}
\item The theoretical distance modulus is
\begin{equation}
\mu(Host, Cosmology) = \mu_{wCDM}(z; \Omega_M, w)
\end{equation}
is the predicted distance modulus of a flat dark energy cosmology
with parameters $\Omega_M$ and $w$, and host redshift $z$.

The flux incident on Earth is the redshifted and time-dilated SED scaled
by the distance modulus
\begin{equation}
f(Luminosity, Host, \mu) = 10^{-\mu/2.5} Luminosity(\lambda/(1+z)).
\end{equation}

The expected photometric signal in counts depends on the flux and throughput of the
optical chain.  A first model encapsulates flux calibration through a zeropoint
\begin{equation}
counts(Throughput, f) = 10^{Z/2.5} \int  f(\lambda) T_0(\lambda)d\lambda,
\end{equation}
where $T_0$ is a fiducial transmission function for the appropriate filter.  A more sophisticated
analysis can have a transmission function specific to each observation.

Practically the photometric extraction is handled by independent software, which calculates
the $n$ with uncertainties $C_P$
by modeling and separating out polluting background light from the transient
signal
\begin{equation}
Photometry_o |  Luminosity, Cosmology,Host,Throughput \sim \mathcal{N}(counts,C_P).
\end{equation}
Dillon is providing this.
Note that the model has explicit calibration parameters that are to be fit.  An alternative
approach is to include calibration uncertainties in $C_P$.

Recall that the sample likelihood has terms
$\epsilon(Photometry_o, b_k)p(data| model)$, which turn out to be truncated Gaussians.

\item Spectroscopic measurements are assumed
to provide prefect determinations of transient type
\begin{equation}
Type_o | Type  \sim \delta(Type-Type_o).
\end{equation}
\item There are measurements of the host galaxy properties.
The first complication is
the correct identification of the host.  For transients that have spectroscopic
confirmation, an unambiguous host determination is assumed such that
\begin{align}
Redshift_o  | Host & \sim \delta(Redshift_o- z)\\
Host~properties_o  | Host & \sim \mathcal{N}(p, C),
\end{align}
where $p$ are the Host properties being measured,
$C$ is the covariance of the measurement of the galaxy identified as the host.
DES plans to get spectroscopic redshifts with negligible measurement uncertainties on all
possible hosts.

For the photometric sample, there is uncertainty in the identity of the host galaxy.
From a set of potential host galaxies $\hat{G}$ each with probability $\pi$ of being the correct host,
\begin{equation}
G \sim \prod_i \delta(G-\hat{G}_i) \pi_i.
\end{equation}
Ravi can provide for each transient a list of candidate hosts and their probabilities.
Thus for the photometric sample
\begin{align}
Redshift_o  | Host &  \sim \sum_i \pi_i \delta(Redshift_o- z_i)\\
Host~properties_o  | Host & \sim  \sum_i \pi_i \mathcal{N}(p_i, C),
\end{align}
\end{itemize}

This accounts for all the terms needed to assemble to pdf of the underlying distribution,
from which the likelihood of the analysis sample is assembled.

@@

\subsubsection{Likelihood for the {\it obs} Set}
%\begin{multline}
%p(n_{SN}, G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
%= \sum_i \int_L dL\, p(n_{SN}, G | b_i, L, \ldots)
%P(b_{i} |L, \ldots) P(L|\ldots) ,
%\end{multline} 
%\begin{equation}
%p(n_{SN}, G | b_i, L, \ldots)
%  =  \frac{ \prod_j
% \epsilon(n_{SN,j}; b_{ij})
%  p(n_{SN}, G | L, \ldots)}
%  {\prod_j P(b_{ij}{ |L,\dots)}}.
%\label{misconf:eqn}
%\end{equation}
